---
format:
  revealjs:
    theme: night
    embed-resources: true
    slide-number: true
    navigation-mode: linear
    controls: true
    css: styles.css
    fontsize: "25pt"
---

# Referencing {background-color="#FFCD00"}

## BibTex Keys

Now that we know how to write in Markdown, let's add some references!

Adding reproducible references happens with **BibTex keys**. A typical BibTex key might look like as follows:

``` markdown
@article{nash51,
  author  = "Nash, John",
  title   = "Non-cooperative Games",
  journal = "Annals of Mathematics",
  year    = 1951,
  volume  = "54",
  number  = "2",
  pages   = "286--295"
}
```

## Working With BibTex Keys

We will work with with bibliographies in the form of `.bib` files (*BibTeX Bibliographical Database*). `.bib` files are text files which contain a list of references in the form of BibTex keys.

Here is an example of another BibTex key for a reference used in this workshop:

``` markdown
@misc{RMarkdownWritingReproducible,
    title = {{RMarkdown} for writing reproducible scientific papers},
    url = {https://libscie.github.io/rmarkdown-workshop/handout.html},
    urldate = {2023-04-18},
}
```

## Working With BibTex Keys {.smaller}

The typical workflow is as follows:

::: incremental
1.  Find a reference that you need in your manuscript.
2.  Save it to your Zotero library via the browser extension\
    Zotero automatically updates your `.bib` file through syncing with Zotero desktop and updating the `.bib` file.
3.  Cite the references using the `@` + reference identifier:
    -   **In-text citations**: `@nash51` OR `@RMarkdownWritingReproducible`
    -   **Bracketed citations**: `[@nash51]` OR `[@nash51, @RMarkdownWritingReproducible]`
:::

. . .

``` yaml
---
bibliography: references.bib
---
```

More info: <https://quarto.org/docs/authoring/footnotes-and-citations.html>

## References on the fly

-   Alternatively, when using the Visual Editor option - you can use the citation dialog/option to insert citations directly.

-   Let's go to RStudio!

## Exercise

Go to the *`Referencing`* section on the website and refer to the instructions in the *Exercise* section.

You will go through the workflow of creating a `.bib` file that will be automatically updated by Zotero and integrating it with your Quarto document.

```{r}
library(countdown)

countdown(minutes = 10,
          color_background = "#FFCD00")
```

## APA Style

-   You need to use a csl (citation style guide) file to make sure references are properly formatted

-   I have included the apa7.csl file in the project folder

    -   This will format references/citations in accordance with APA 7th edition

## Exercise

-   Add the `csl` to your YAML and include the apa7.csl file

    ```{r}
    #| eval: false
    csl: apa7.csl
    ```

-   Render your document

```{r}

countdown(minutes = 2,
          color_background = "#FFCD00")
```
